---
title: "portland demo"
author: "Taylor Holdaway"
date: "4/8/2016"
output: html_document
---
---
title: "pdx demo"
output: html_document
---
```{r, warning=FALSE, echo=FALSE, message=FALSE}
library(ggplot2)
library(reshape2)
library(ggmap)
library(dplyr)
library(broom)
library(maptools)
library(rgeos)
library(sp)
library(plm)
library(pglm)
```

First we need to import, clean, and geocode the address data. Save it as a csv for future use. 
```{r geocode and clean data, eval=FALSE, message=FALSE, warning=FALSE}
pdx <- read.csv("pdx dmo.csv", header = F)
pdx$V1 <- NULL
pdx$V2 <- NULL
names(pdx) <- c("address","year")
pdx$address <- tolower(as.character(pdx$address))

pdx$year[is.na(pdx$year)] <- 2014

coord <- ggmap::geocode(pdx$address,messaging = F, output = "latlona")
demo.address <- merge(pdx,coord, by = 0)
demo.address$address.x <- NULL
demo.address$Row.names <- NULL
demo.address <- rename(demo.address, address = address.y)
demo.address$address <- as.character(demo.address$address)

write.csv(demo.address, file = "data/pdx_dmo_ref.csv",row.names = F)

portland <- ggmap::get_map(location="portland OR", source = "stamen", maptype = "toner", crop = F,zoom=12)

#SAVE SAVE SAVE
save(portland, file="data/map_portland.rda")

```

Using a Zillow shapefile, I can overlay neighborhoods onto a map of Portland
```{r Neighborhood Polys, message=FALSE, warning=FALSE}
#path and name of shapefile
file <- "data/ZillowNeighborhoods-OR/ZillowNeighborhoods-OR.shp"
shape <- maptools::readShapePoly(file)
shape <- shape[shape@data$CITY != "Salem" & shape@data$CITY != "Eugene",]
shape.df <- broom::tidy(shape, region = ("REGIONID"))
shape.df$id <- as.numeric(shape.df$id)
rm(file)
#load map file
load("data/map_portland.rda")
#overlay neighborhood shapefile on map of portland
mapNeighborhoods <- ggmap(portland) +
  geom_polygon(aes(x=long,y=lat, group=group), alpha = 0.25, data=shape.df, fill='white') + 
  geom_polygon(aes(x=long,y=lat, group=group), data=shape.df, color='deeppink4', fill=NA)
mapNeighborhoods
```

Then, using the geocoded points for demolition permits, I can overlay demolitions on a map of Portland, with neighborhood boundaries in pink.
```{r Demo Points Overlay, message=FALSE, warning=FALSE}
#read the georef'd address of the demo locations
demo.address <- read.csv("data/pdx_dmo_ref.csv")
#plot demolition permits across portland
mapAddress <- mapNeighborhoods + geom_point(aes(x = lon, y = lat, color = as.factor(demo.address$year), shape = as.factor(demo.address$year)), data = demo.address, alpha = 0.8, size = 2.5) + scale_color_hue("Year") + scale_size(guide = "none") + scale_shape(guide = "none")
mapAddress
```

I can then sum the number of demolition permits in each neighborhood polygon. The polygons are then collored according to this number.

```{r Demo Counts Polys, message=FALSE, warning=FALSE}
coordinates(demo.address) <- ~lon+lat
#overlay the count data with the shapefile

data <- over(subset(demo.address, year == 2014), shape[,c("NAME","REGIONID")]) %>%
  dplyr::mutate(year = 2014)
data <- over(subset(demo.address, year == 2015), shape[,c("NAME","REGIONID")]) %>%
  dplyr::mutate(year = 2015) %>%
  rbind(data) %>% 
  na.omit()

#stack up shape.df (two copies, one for each year
shape.df$year <- 2014
shape.df2 <- shape.df %>%
  dplyr::mutate(year = 2015) %>%
  rbind(shape.df)
  
#merge counts with the polygone dataframe
shape.democounts.df <- 
#count number of points in each polygon
  data %>% group_by(REGIONID, year) %>% dplyr::summarize(count = n()) %>%
  dplyr::right_join(shape.df2,by = c("REGIONID"="id", "year"="year")) %>%
  dplyr::arrange(order) %>%
#recode missing values as zero
  dplyr::mutate(count = replace(count,is.na(count),0))

#save counts for later separately:
counts <- data %>% group_by(REGIONID, NAME, year) %>% dplyr::summarize(count = n())
#clean up
rm(data)


#get back the old points dataframe, so that we can plot the points on the same plot. must do this because the old one was turned into a spatial points class
demo.address <- read.csv("data/pdx_dmo_ref.csv")

#plot the polygons shaded by number of demolitions in that neighborhood
mapCounts.2014 <- ggmap(portland) + 
  geom_polygon(aes(x=long,y=lat, group=group, fill=count), data=subset(shape.democounts.df,year==2014), color='black') +
  scale_fill_gradient(low='white', high='red')
#with the demo points indicated, to check for correctness
#  geom_point(aes(x = lon, y = lat, shape = as.factor(demo.address$year)), data = demo.address, alpha = 0.4, size = 2.5) + 
#  scale_size(guide = "none") + scale_shape("Year")
#plot
mapCounts.2014

mapCounts.2015 <- ggmap(portland) + 
  geom_polygon(aes(x=long,y=lat, group=group, fill=count), data=subset(shape.democounts.df,year==2015), color='black') +
  scale_fill_gradient(low='white', high='red')
#with the demo points indicated, to check for correctness
#  geom_point(aes(x = lon, y = lat, shape = as.factor(demo.address$year)), data = demo.address, alpha = 0.4, size = 2.5) + 
#  scale_size(guide = "none") + scale_shape("Year")
#plot
mapCounts.2015
```

Also using data from Zillow, I can color the neighborhood polygons according to median(?) house price. I also show house price from 2014 and 2013.
```{r Zillow Prices, message=FALSE, warning=FALSE}

prices <- read.csv("data/ZillowPrices/portland-or-neighborhoods-Zillow-Home-Value-Index-TimeSeries.csv",stringsAsFactors = F)
#lowercase
names(prices) <- tolower(names(prices))
#format prices as numbers (get rid of $ and ,)
prices[,4:123] <- (gsub("\\$", "", as.matrix(prices[,4:123])))
prices[,4:123] <- as.numeric(gsub(",","", as.matrix(prices[,4:123]) ))
#reshape to long, by date
prices <- melt(prices,variable.name="date",value.name="price")
prices <- colsplit(as.character(prices$date),"[.]",c("month","year")) %>%
  cbind( . , prices)

#calculate yearly averages for each neighborhood
prices <- prices %>%
  select(year,region.name,price) %>%
  group_by(region.name,year) %>%
  summarize(avg_price = mean(price))
#now, merge with the polygon dataframe (but first we make a new such polygon because the old one had the REGIONID variable as )
shape.prices.df <- shape %>% tidy(region="NAME") %>%
  dplyr::left_join(prices,by = c("id"="region.name")) %>%
  arrange(order)

demo.address.2015 <- subset(demo.address, year == 2015)
demo.address.2014 <- subset(demo.address, year == 2014)


mapPrices.2014 <- ggmap(portland) + geom_polygon(aes(x=long,y=lat, group=group, fill=avg_price), alpha = 0.8, data=subset(shape.prices.df, year==2014), color='black') + scale_fill_gradient("Average Zillow Price Index",low='blue', high='red', limits=c(min(shape.prices.df$avg_price,na.rm=T),max(shape.prices.df$avg_price,na.rm=T))) +
  geom_point(aes(x = lon, y = lat), data = subset(demo.address, year==2014), alpha = 0.4, size = 2.5) + scale_size(guide = "none") + scale_shape("Year") + ggtitle("Demolitions and House Prices for 2014")
mapPrices.2014

mapPrices.2015 <- ggmap(portland) + geom_polygon(aes(x=long,y=lat, group=group, fill=avg_price), alpha = 0.8, data=subset(shape.prices.df, year==2015), color='black') + scale_fill_gradient("Average Zillow Price Index",low='blue', high='red', limits=c(min(shape.prices.df$avg_price,na.rm=T),max(shape.prices.df$avg_price,na.rm=T))) +
  geom_point(aes(x = lon, y = lat), data = subset(demo.address, year==2015), alpha = 0.4, size = 2.5) + scale_size(guide = "none") + scale_shape("Year") + ggtitle("Demolitions and House Prices for 2015")
mapPrices.2015
```

I then run a regression using OLS of the number of demolition permits in each neighborhood on the median house price, to test for correlation between the prices of houses in a neighborhood and the number of demo permits issued there. I also look for correlation between median neighborhood house price in year $t-1$ and demo permits issued in year $t$.
```{r}

counts.prices.df <- prices %>% 
  dplyr::filter(year %in% c(2012:2015)) %>%
  left_join(counts,by=c("region.name"="NAME", "year"="year")) %>%
  dplyr::mutate(count = replace(count,is.na(count) & year %in% c(2014,2015),0)) %>%
  dplyr::select(-REGIONID)

counts.prices.pdf <- pdata.frame(counts.prices.df, index = c("region.name", "year"), drop.index = F, row.names = T)

reg <- lm(count~avg_price+lag(avg_price,k=1),data = counts.prices.pdf)
reg1 <- plm(count~avg_price+lag(avg_price,k=1:2), data = counts.prices.pdf, model = "within", effect = "twoway")
summary(reg)
summary(reg1)
poisson <- pglm
```

I fit a regression tree to examine the possibility of nonlinear relationships in the data.
```{r}

```

I also examine the possibility of other important clustering in the data (how do I do this!?!?)

See how many clusters are there (scree plot) across years. More clusters means more dispersed geographically.
```{r}

```



http://www.r-bloggers.com/plotting-choropleths-from-shapefiles-in-r-with-ggmap-toronto-neighbourhoods-by-population/

http://www.r-bloggers.com/shapefile-polygons-plotted-on-google-maps-using-ggmap-in-r-throw-some-throw-some-stats-on-that-mappart-2/